Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-19 paper/paper.tex
Update DPDS description

The static provisioning approach of SPDS may not be optimal when resource demand
changes during ensemble execution. In those cases, SPDS may either over-provision, 
wasting the budget on idle VMs, or under-provision, using up the deadline while starving 
the application of usable VMs. One of the advantages of elasticity in cloud computing
is that the number of resources can be changed in response to application demand. 
In policy-based provisioning (or auto-scaling) systems, the number of provisioned VMs 
is adjusted based on application demand as measured by metrics such as resource 
utilization and queue length. The DPDS algorithm uses resource utilization
to determine when to provision and de-provision VMs. The provisioner
module periodically computes utilization as the percentage of idle VMs. The
provisioner algorithm detects which VMs are finishing their hourly billing cycle
during the next provisioning period and decides whether they should continue
running, should be terminated or new VMs should be provisioned: see
Algorithm~\ref{alg:prov}.

Because it is assumed that VMs are billed by the hour, DPDS only considers VMs that are
approaching their hourly billing cycle when deciding which VMs to terminate.

\begin{algorithm}
\caption{Dynamic provisioning algorithm for DPDS}
\label{alg:prov}
\begin{algorithmic}[1]
\Require $c$: consumed budget; $b$: total budget; $d$: deadline; $p$: price;
$t$: current time; $u_h$: upper utilization threshold; $u_l$: lower utilization
threshold.
\Procedure{Provision}{}
    \State $V_C\gets set\ of\ VMs\ completing\ billing\ cycle$
    \State $n_T\gets\ 0$ \Comment{number of VMs to terminate} 
    \If{$ b - c < p * |V_C| \vee t > d $ }
    	\State $V_R\gets set\ of\ running\ VMs$
    	\State $n_T\gets |V_R| - \lfloor(b-c)/p\rfloor$ 
    	\State $V_T\gets select\ n_T VMs\ to\ terminate\ from\ V_C$
    	\State \Call{Terminate}{$V_T$}
    \Else 
    	\If{$u>u_h$}
    		\State \Call{Start}{$new\ VM$}
    	\EndIf
    	\If{$u<u_l$}
    		\State $V_F\gets free\ (idle)\ VMs$
    		\State $n_T\gets \lceil|V_F|/2\rceil$ 
    		\State \Call{Terminate}{$V_T$}
    	\EndIf 
    \EndIf
\EndProcedure
\end{algorithmic} 
\end{algorithm}

The number of instances to terminate $n_T$ computed in line 6 is the number that
would overrun the budget if not terminated. On the other hand $n_T$ computed in
line 15 prevents terminating all the instances too quickly, but makes sure that
if there is only one instance left it should be terminated. However, the
termination process is not immediate: to make sure that VMs that have been
already paid for are not terminated prematurely the \textsc{Terminate} procedure
terminates only these instances that are close to the end of their billing
cycle. The remaining time threshold is calculated as a sum of provisioner
interval and estimated termination delay. This guarantees that no VMs can
overrun the assigned budget. To avoid uncontrolled increase of number of
instances, which may happen in the case of highly parallel workflows,
provisioner checks also the maximum scaling factor from the initial number of
provisioned VMs and prohibits starting new ones when this threshold is reached.

>>>>

The static provisioning approach of SPDS may not be optimal when resource demand
changes during ensemble execution. In those cases, SPDS may either over-provision, 
wasting the budget on idle VMs, or under-provision, using up the deadline while starving 
the application of usable VMs. One of the advantages of elasticity in cloud computing 
is that the number of resources can be changed to meet the demands of an application.
In policy-based provisioning (or auto-scaling) systems, metrics such as resource 
utilization and queue length are monitored to estimate application demand. The DPDS 
algorithm uses this approach to determine when to provision and de-provision VMs. The 
algorithm periodically computes resource utilization as the percentage of idle VMs, and
adjusts the number of VMs if the utilization is above or below given thresholds.
Because it is assumed that VMs are billed by the hour, DPDS only considers VMs that are
approaching their hourly billing cycle when deciding which VMs to terminate. This process
is described in Algorithm~\ref{alg:prov}.

\begin{algorithm}
\caption{Dynamic provisioning algorithm for DPDS}
\label{alg:prov}
\begin{algorithmic}[1]
\Require $c$: consumed budget; $b$: total budget; $d$: deadline; $p$: price;
$t$: current time; $u_h$: upper utilization threshold; $u_l$: lower utilization
threshold; $v_{max}$: maximum number of VMs
\Procedure{Provision}{}
	\State $V_R\gets$ set of running VMs
    \State $V_C\gets$ set of VMs completing billing cycle
    \If{$ b - c < |V_C| * p\ or\ t > d $ }
    	\State $n_T\gets |V_R| - \lfloor(b-c)/p\rfloor$
    	\State $V_T\gets$ select $n_T$ VMs to terminate from $V_C$
    	\State \Call{Terminate}{$V_T$} \label{l:terminate1}
    \Else 
		\State $u\gets$ current VM utilization
    	\If{$u>u_h$ and $|V_R| < v_{max}$}
    		\State \Call{Start}{$new\ VM$}
    	\ElsIf{$u<u_l$}
    		\State $V_I\gets$ set of idle VMs
    		\State $n_T\gets \lceil|V_I|/2\rceil$ \label{l:nT2}
			\State $V_T\gets$ select $n_T$ VMs to terminate from $V_I$
    		\State \Call{Terminate}{$V_T$} \label{l:terminate2}
    	\EndIf 
    \EndIf
\EndProcedure
\end{algorithmic} 
\end{algorithm}

The set of VMs completing their billing cycle is determined by considering both the
provisioner interval, and the termination delay of the provider. This guarantees 
that VMs can be terminated before they start the next billing cycle and prevents the
budget from being overrun. The VMs terminated in line \ref{l:terminate1} are the ones 
that would overrun the budget if not terminated in the current provisioning cycle. 
The VMs terminated on line \ref{l:nT2} are chosen to increase the resource utilization
to the desired threshold. In order to prevent instances from being
terminated too quickly, potentially wasting resources that have already been paid for
but could be used later, no more than half of the idle resources are terminated during
each provisioning cycle. To avoid an uncontrolled increase in the number of
instances, which may happen in the case of highly parallel workflows, the provisioner 
will not start a new VM if the user-defined scaling parameter, $v_{max}$, has
been reached.
